Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-08-07 paper/paper.tex
added a whole bunch of things and edited Makefile so can pull and make paper Tested everything on node21.

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-300-50-0-0-1500.pdf}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/single-image-stoch}
  \caption{Countour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic simulation and the lower shows results from the
    stochastic simulation.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space as guassians with
    characteristic length equal to the Abbe diffraction limit of
    $600nm$ wavelength light}
  \label{image-p}
\end{figure*}

@@

\begin{figure}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  The vertical axis shows stacked and changing over
    time the total number of proteins that are of four different
    compound stages and in the different sections of the cell.  Both
    shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-stad/plots/arrow-400-stoch}
  \caption{Arrow plots of our stadium shape}
  \label{stadium-arrow}
\end{figure}

@@

\section{Our '\textgreater' Shaped Cells}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-525-525-9900-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-550-550-9900-1500}
  \label{fig:Vbox}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9900-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1400-1400-9900-1500}
  \label{fig:Vbox2}
\end{figure}

We have also simulated a flattened '\textgreater' shape that shows an
interesting size dependent pattern (seen in
Figures~\ref{fig:Vbox}~and~\ref{fig:Vbox2}).  When the shape is small enough
that the distance from the apex tip to the opposite curve is less than
our stability limit, a simple, two pole oscillation pattern develops
and is sustained (where the two poles are on opposite leftward corners
of the cell).  When the cell is increased in size so that this
distance is approximately equal to our stability value, the cell
develops a double oscillation in time, where the maxima oscillates
from the lower corner to the rightward apex, then \emph{back to the
  lower corner}, then to the apex again, then to the upper corner,
then back to the apex, then to the upper corner, then to the apex
again, then back to the lower corner, and continues this pattern
indefinately.  Figure~\ref{fig:Vbox} shows the smaller size, for which the
protein oscillates simply, and an intermediate size, which shows a
pattern for which one out of three spikes in the regular oscillation
is less significant.  This second pattern seems to be a precursor to
the double oscillation that seen in Figure~\ref{fig:Vbox2}.  The second
plot shown in Figure~\ref{fig:Vbox2} shows a slightly larger cell for
which the distance measured is larger than our stability limit
distance.  This simulation shows a random pattern.

A detailed analysis of a time series of protein contour plots shows
that the importance of spatial dimensions is based on the ability of
protein in different stages to diffuse neccesary distances before
undergowing certain types of reactions.  The double oscillation
pattern shows that every time the membrane bound MinD maximizes in the
upper corner as opposed to the rightward apex, it does so because the
MinE has not had enough time to diffuse to this upper corner, leaving
a MinE free zone.  This particular pattern is specific to this shape
and size, but in general its the restrictions set in place by the
rates at which proteins diffuse and the distances they cover before
reacting that dictate their oscillatory bahavoir.

Simulations of this shape in which the two planar dimensions are large
compared to those shown(roughly $9\micron$ across) show no stable
regular periodic pattern. Animations show that the MinD acts as a sort
of traveling cluster that is constantly being eaten away by MinE
behind it.  When this cluster enters a region of high curvature that
ends in a dead end corner it performs the polar maxima behavoir seen
in the other simulations, regrouping out away from the corner where it
travels in the opposite direction while being ``chased'' away by the
MinE that itself is currently maximizing in the corner.  The maxima
occur on non-polar walls often, and will often split into two roaming
clusters.

>>>>

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-300-50-0-0-1500-exact}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/single-image-plot--p-300-50-0-0-1500-full_array}
  \caption{Countour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic simulation and the lower shows results from the
    stochastic simulation.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space as guassians with
    characteristic length equal to the Abbe diffraction limit of
    $600nm$ wavelength light (ends up being about 200nm)\fixme{say
      something about the axial resolution, which is even more limited
      than the lateral resolution, which is why we integrate data long
      axial direction}}
  \label{image-p}
\end{figure*}

@@

\begin{figure}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500-exact}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500-exact}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  The vertical axis shows stacked and changing over
    time the total number of proteins that are of four different
    compound stages and in the different sections of the cell.  Both
    shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-stad/plots/ave-time/plot-time-averaged-arrow-300-1600-NflD-stad-25-400-150-0-1500-full_array}
  \caption{Arrow plots of our stadium shape}
  \label{stadium-arrow}
\end{figure}

@@

%% \section{Our '\textgreater' Shaped Cells}
%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-525-525-9900-1500}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-550-550-9900-1500}
%%   \label{fig:Vbox}
%% \end{figure}
%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9900-1500}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1400-1400-9900-1500}
%%   \label{fig:Vbox2}
%% \end{figure}

%% We have also simulated a flattened '\textgreater' shape that shows an
%% interesting size dependent pattern (seen in
%% Figures~\ref{fig:Vbox}~and~\ref{fig:Vbox2}).  When the shape is small enough
%% that the distance from the apex tip to the opposite curve is less than
%% our stability limit, a simple, two pole oscillation pattern develops
%% and is sustained (where the two poles are on opposite leftward corners
%% of the cell).  When the cell is increased in size so that this
%% distance is approximately equal to our stability value, the cell
%% develops a double oscillation in time, where the maxima oscillates
%% from the lower corner to the rightward apex, then \emph{back to the
%%   lower corner}, then to the apex again, then to the upper corner,
%% then back to the apex, then to the upper corner, then to the apex
%% again, then back to the lower corner, and continues this pattern
%% indefinately.  Figure~\ref{fig:Vbox} shows the smaller size, for which the
%% protein oscillates simply, and an intermediate size, which shows a
%% pattern for which one out of three spikes in the regular oscillation
%% is less significant.  This second pattern seems to be a precursor to
%% the double oscillation that seen in Figure~\ref{fig:Vbox2}.  The second
%% plot shown in Figure~\ref{fig:Vbox2} shows a slightly larger cell for
%% which the distance measured is larger than our stability limit
%% distance.  This simulation shows a random pattern.

%% A detailed analysis of a time series of protein contour plots shows
%% that the importance of spatial dimensions is based on the ability of
%% protein in different stages to diffuse neccesary distances before
%% undergowing certain types of reactions.  The double oscillation
%% pattern shows that every time the membrane bound MinD maximizes in the
%% upper corner as opposed to the rightward apex, it does so because the
%% MinE has not had enough time to diffuse to this upper corner, leaving
%% a MinE free zone.  This particular pattern is specific to this shape
%% and size, but in general its the restrictions set in place by the
%% rates at which proteins diffuse and the distances they cover before
%% reacting that dictate their oscillatory bahavoir.

%% Simulations of this shape in which the two planar dimensions are large
%% compared to those shown(roughly $9\micron$ across) show no stable
%% regular periodic pattern. Animations show that the MinD acts as a sort
%% of traveling cluster that is constantly being eaten away by MinE
%% behind it.  When this cluster enters a region of high curvature that
%% ends in a dead end corner it performs the polar maxima behavoir seen
%% in the other simulations, regrouping out away from the corner where it
%% travels in the opposite direction while being ``chased'' away by the
%% MinE that itself is currently maximizing in the corner.  The maxima
%% occur on non-polar walls often, and will often split into two roaming
%% clusters.
